Perturbing SU($n$)-symmetric interactions

### 2020-03-10 collective_sun/sun_perturbation.tex
explicitly exclude the empty set in the domain of diagrams

To simplify the product in \eqref{eq:sym_prod_proj}, we first write
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}}
  w_j\p{k} O_j\p{k},
  \label{eq:sym_prod_vec}
\end{align}
where we extend the definitions of $w_j\p{k}$ and $O_j\p{k}$ to all
vectors $k\in\ZZ_N^{M_j}$ by defining symmetric dimension-$M_j$
tensors $w_j$ and $O_j$ (i.e.~tensors with $M_j$ indices, with $O_j$
an {\it operator-valued} tensor) for which $w_j\p{k}=O_j\p{k}=0$ if
$k$ contains any repeated symbols.  We then collect terms whose
operator content is equivalent up to a permutation of spins.  To this
end, we classify terms in \eqref{eq:sym_prod_vec} by the numbers $g_S$
of indices shared by all tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.
For example, a term of the form
$w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
$\p{a,b,c,d,e}\in\C_N\p{5}$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,3}} = 1,
  &&
  g_{\set{2,3}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
assignment can be represented by the Venn diagram
\begin{align}
  \diagram{example_123},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of
circles $\set{w_j:j\in S}$.  We denote an assignment of distinct
indices according to choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ by
$g$, where $\PP\p{\ZZ_p}$ is the power set (i.e.~set of all subsets)
of $\ZZ_p$.  Given a vector $\m M$ of the dimensions $M_j$, we then
denote the set of all valid index assignments $g$ by $\G\p{\m M}$.
For consistency, all valid index assignments $g\in\G\p{\m M}$ must
assign exactly $M_j$ indices to the dimension-$M_j$ tensor $w_j$,
i.e.~
\begin{align}
  \sum_{S\in\PP\p{\ZZ_p}\,:\,j\in S} g_S
  = \sum_{R\in\PP\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
  = M_j
\end{align}
for all $j\in\ZZ_p$.  In order to make the index assignments
determined by choices of $g_S$ for all $S\in\PP\p{\ZZ_p}$ unique, we
also enforce that $g_{\set{}}=0$ for all $g\in\G\p{\m M}$, which is
equivalent to saying that every dot in the corresponding Venn diagram
must lie within some circle.  The set of valid index assignments
$\G\p{\m M}$ is then essentially the set of all Venn diagrams of the
form in \eqref{eq:venn_diagram} with $p=\dim\p{\m M}$ circles and a
total of $M_j$ dots within circle $w_j$.

@@

In order to write out the factors in \eqref{eq:sym_prod_group_start}
explicitly, we identify the set of values that the distinct indices
assigned according to $g$ can take:
\begin{align}
  \C_N\p{g}
  \equiv \set{ k \in \bigotimes_{S\in\PP\p{\ZZ_p}} \C_N\p{g_S}
    : \t{all elements of $k$ are distinct} },
  \label{eq:index_values}
\end{align}
where each $k\in\C_N\p{g}$ has the natural decomposition
$k=\p{k_S:S\in\PP\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  In words,
$\C_N\p{g}$ consists of all the ways to choose, for each subset $S$ of
the tensors $\set{w_j}$, a set $k_S$ of $g_S$ distinct spins, with the
restriction that all $k_S$ are orthogonal: if $z\in k_S$, then
$z\notin k_R$ for $R\ne S$.  Having defined the set $\C_N\p{g}$ of
index values, we define the restriction of $k\in\C_N\p{g}$ to values
associated with $w_j$ and $O_j$:
\begin{align}
  k_j \equiv \bigcup_{S\in\PP\p{\ZZ_p} : j\in S} k_S
  = \bigcup_{R\in\PP\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
\end{align}
where the union of lists $k_S$ denotes a concatenation of those lists.
These definitions allow us to expand
\begin{align}
  w\p{g} \equiv \sum_{k\in\C_N\p{g}} \prod_{j\in\ZZ_p} w_j\p{k_j},
  &&
  O\p{g} \equiv \prod_{j\in\ZZ_p}
  O_j\p{\ell_j}~\t{for some}~\ell\in\C_N\p{g},
  \label{eq:diagram_factors}
\end{align}
where the choice of $\ell\in\C_N\p{g}$ does not matter after
projection onto permutationally symmetric manifold, and
\begin{align}
  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} \C\p{ g_S : j\in S }}
  \sp{\prod_{S\in\PP\p{\ZZ_p}} \p{g_S!}^\abs{S}}
  = \prod_{j\in\ZZ_p} M_j!,
  \label{eq:sym_prod_symmetry}
\end{align}
where the first product in \eqref{eq:sym_prod_symmetry} contains
multinomial coefficients
\begin{align}
  \C\p{a_1,a_2,\cdots,a_m}
  \equiv \f{\p{a_1+a_2+\cdots+a_m}!}{a_1!a_2!\cdots a_m!}
\end{align}
that account for the number of ways to partition the indices of each
$w_j,O_j$ into sets of shared indices as specified by $g$, and the
second product in \eqref{eq:sym_prod_symmetry} accounts for the number
of ways to permute each set of shared indices on every tensor.
Altogether, the symmetry factors $\S\p{g}$ cancel out with the
$\sim M_j!$ prefactors in \eqref{eq:sym_prod_vec}, so
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  \EQPS \sum_{g\in\G\p{\m M}} w\p{g} O\p{g}.
  \label{eq:sym_prod_group}
\end{align}

@@

The simplest diagrams have $g_S$ filled dots (``$\bullet$'') at the
intersection of the circles $j\in S\in\PP\p{\ZZ_p}$, and are simply
equal to the associated scalar $w\p{g}$ defined in
\eqref{eq:diagram_factors}, e.g.
\begin{align}
  \diagram{example_123}
  \equiv \sum_{\substack{a,b,c,\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
      \abs{\set{a,b,c,d,e}}=5}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
\end{align}
where the restriction $\abs{\set{a,b,c,d,e}}=5$ denotes that all of
$a,b,c,d,e$ are distinct.  Such a diagram with $\abs{g}$ dots
nominally takes $\O\p{N^{\abs{g}}}$ time to compute.  In practice,
this computational cost can be prohibitive for performing numerical
simulations of systems with any appreciable size.  We therefore wish
to simplify diagrams in order to reduce their computational cost as
much as possible.  Our general strategy will be to replace constrained
sums that make all indices interdependent by unconstrained sums that
can be carried out independently, in sequence.  In order to represent
different constraints, we need to define diagrams in which filled dots
(``$\bullet$'') may be replaced by empty dots (``$\circ$'') or crosses
(``$\bm\times$'').

>>>>

To simplify the product in \eqref{eq:sym_prod_proj}, we first write
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  = \prod_{j\in\ZZ_p} \f1{M_j!} \sum_{k\in\ZZ_N^{M_j}}
  w_j\p{k} O_j\p{k},
  \label{eq:sym_prod_vec}
\end{align}
where we extend the definitions of $w_j\p{k}$ and $O_j\p{k}$ to all
vectors $k\in\ZZ_N^{M_j}$ by defining symmetric dimension-$M_j$
tensors $w_j$ and $O_j$ (i.e.~tensors with $M_j$ indices, with $O_j$
an {\it operator-valued} tensor) for which $w_j\p{k}=O_j\p{k}=0$ if
$k$ contains any repeated symbols.  We then collect terms whose
operator content is equivalent up to a permutation of spins.  To this
end, we classify terms in \eqref{eq:sym_prod_vec} by the numbers $g_S$
of indices shared by all tensors $w_j,O_j$ with $j\in S\subset\ZZ_p$.
For example, a term of the form
$w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}$ with distinct indices
$\p{a,b,c,d,e}\in\C_N\p{5}$ would have
\begin{align}
  g_{\set{1}} = 1,
  &&
  g_{\set{1,2,3}} = 1,
  &&
  g_{\set{1,3}} = 1,
  &&
  g_{\set{2,3}} = 2,
\end{align}
and $g_S=0$ for all other subsets $S\subset\ZZ_3$.  This index
assignment can be represented by the Venn diagram
\begin{align}
  \diagram{example_123},
  \label{eq:venn_diagram}
\end{align}
where $g_S$ is determined the number of dots at the intersection of
circles $\set{w_j:j\in S}$.  Denoting the set of all nonempty subsets
of $\ZZ_p$ by $\PP^*\p{\ZZ_p}$, we denote an assignment of distinct
indices according to choices of $g_S$ for all $S\in\PP^*\p{\ZZ_p}$ by
$g$.  Given a vector $\m M$ of the dimensions $M_j$, we then denote
the set of all valid index assignments $g$ by $\G\p{\m M}$.  For
consistency, all valid index assignments $g\in\G\p{\m M}$ must assign
exactly $M_j$ indices to the dimension-$M_j$ tensor $w_j$, i.e.~
\begin{align}
  \sum_{S\in\PP^*\p{\ZZ_p}\,:\,j\in S} g_S
  = \sum_{R\in\PP^*\p{\ZZ_p\setminus\set{j}}} g_{\set{j}\cup R}
  = M_j
\end{align}
for all $j\in\ZZ_p$.  The set of valid index assignments $\G\p{\m M}$
is then essentially the set of all Venn diagrams of the form in
\eqref{eq:venn_diagram} with $p=\dim\p{\m M}$ circles and a total of
$M_j$ dots within circle $w_j$.

@@

In order to write out the factors in \eqref{eq:sym_prod_group_start}
explicitly, we identify the set of values that the distinct indices
assigned according to $g$ can take:
\begin{align}
  \C_N\p{g}
  \equiv \set{ k \in \bigotimes_{S\in\PP^*\p{\ZZ_p}} \C_N\p{g_S}
    : \t{all elements of $k$ are distinct} },
  \label{eq:index_values}
\end{align}
where each $k\in\C_N\p{g}$ has the natural decomposition
$k=\p{k_S:S\in\PP^*\p{\ZZ_p}}$ with $k_S\in\C_N\p{g_S}$.  In words,
$\C_N\p{g}$ consists of all the ways to choose, for each subset $S$ of
the tensors $\set{w_j}$, a set $k_S$ of $g_S$ distinct spins, with the
restriction that all $k_S$ are orthogonal: if $z\in k_S$, then
$z\notin k_R$ for $R\ne S$.  Having defined the set $\C_N\p{g}$ of
index values, we define the restriction of $k\in\C_N\p{g}$ to values
associated with $w_j$ and $O_j$:
\begin{align}
  k_j \equiv \bigcup_{S\in\PP^*\p{\ZZ_p} : j\in S} k_S
  = \bigcup_{R\in\PP^*\p{\ZZ_p\setminus\set{j}}} k_{\set{j}\cup R},
\end{align}
where the union of lists $k_S$ denotes a concatenation of those lists.
These definitions allow us to expand
\begin{align}
  w\p{g} \equiv \sum_{k\in\C_N\p{g}} \prod_{j\in\ZZ_p} w_j\p{k_j},
  &&
  O\p{g} \equiv \prod_{j\in\ZZ_p}
  O_j\p{\ell_j}~\t{for some}~\ell\in\C_N\p{g},
  \label{eq:diagram_factors}
\end{align}
where the choice of $\ell\in\C_N\p{g}$ does not matter after
projection onto permutationally symmetric manifold, and
\begin{align}
  \S\p{g} \equiv \sp{\prod_{j\in\ZZ_p} \C\p{ g_S : j\in S }}
  \sp{\prod_{S\in\PP^*\p{\ZZ_p}} \p{g_S!}^\abs{S}}
  = \prod_{j\in\ZZ_p} M_j!,
  \label{eq:sym_prod_symmetry}
\end{align}
where the first product in \eqref{eq:sym_prod_symmetry} contains
multinomial coefficients
\begin{align}
  \C\p{a_1,a_2,\cdots,a_m}
  \equiv \f{\p{a_1+a_2+\cdots+a_m}!}{a_1!a_2!\cdots a_m!}
\end{align}
that account for the number of ways to partition the indices of each
$w_j,O_j$ into sets of shared indices as specified by $g$, and the
second product in \eqref{eq:sym_prod_symmetry} accounts for the number
of ways to permute each set of shared indices on every tensor.
Altogether, the symmetry factors $\S\p{g}$ cancel out with the
$\sim M_j!$ prefactors in \eqref{eq:sym_prod_vec}, so
\begin{align}
  \prod_{j\in\ZZ_p} \sum_{k\in\C_N\p{M_j}} w_j\p{k} O_j\p{k}
  \EQPS \sum_{g\in\G\p{\m M}} w\p{g} O\p{g}.
  \label{eq:sym_prod_group}
\end{align}

@@

The simplest diagrams have $g_S$ filled dots (``$\bullet$'') at the
intersection of the circles $j\in S$, and are simply equal to the
associated scalar $w\p{g}$ defined in \eqref{eq:diagram_factors}, e.g.
\begin{align}
  \diagram{example_123}
  \equiv \sum_{\substack{a,b,c,\in\ZZ_N\\\p{d,e}\in\C_N\p{2}\\
      \abs{\set{a,b,c,d,e}}=5}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e},
\end{align}
where the restriction $\abs{\set{a,b,c,d,e}}=5$ denotes that all of
$a,b,c,d,e$ are distinct.  Such a diagram with $\abs{g}$ dots
nominally takes $\O\p{N^{\abs{g}}}$ time to compute.  In practice,
this computational cost can be prohibitive for performing numerical
simulations of systems with any appreciable size.  We therefore wish
to simplify diagrams in order to reduce their computational cost as
much as possible.  Our general strategy will be to replace constrained
sums that make all indices interdependent by unconstrained sums that
can be carried out independently, in sequence.  In order to represent
different constraints, we need to define diagrams in which filled dots
(``$\bullet$'') may be replaced by empty dots (``$\circ$'') or crosses
(``$\bm\times$'').
